---
title: "Articles Published by US Federal Personnel"
subtitle: "January 2019 - May 2025)"
author: "Emilio M. Bruna"
date: "`r format(Sys.time(), '%d %B %Y')`"
# output: html_document
output: pdf_document
geometry: margin=1in
fontsize: 11pt
linkcolor: darkmidnightblue
urlcolor: darkmidnightblue
header-includes:
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \usepackage[default]{sourcesanspro}
  - \usepackage{parskip}
  - \usepackage{geometry}
  - \usepackage{caption}
  - \usepackage{xcolor}
  - \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
  - \AtBeginDocument{\let\maketitle\relax}
  - \usepackage[none]{hyphenat}
  - \usepackage{setspace}
  - \usepackage[document]{ragged2e}
  - \usepackage{graphicx}
  - \usepackage{sectsty}\allsectionsfont{\raggedright}
  - \usepackage{sectsty}\sectionfont{\centering\color{darkmidnightblue}}
  - \usepackage{sectsty}\subsectionfont{\centering\color{darkmidnightblue}}
  - \usepackage{titlesec}
  - \usepackage[export]{adjustbox}
  - \titlespacing{\section}{0pt}{24pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
  - \titlespacing{\subsection}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
  - \titlespacing{\subsubsection}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
  - \usepackage{floatrow}
  - \DeclareFloatSeparators{mysep}{\hskip-94em}
  # - \floatsetup[figure]{capposition=beside,capbesidesep=mysep,capbesideposition={right, center}}
  - \usepackage{float}
  - \usepackage{wrapfig} 
  - \newcommand{\blandscape}{\begin{landscape}}
  # - \usepackage[font=small,labelfont=bf,labelsep=space]{caption}
  # - \captionsetup{figurename=Fig.,tablename=Table}
  # - \usepackage[labelfont=bf]{caption}
  # - \captionsetup{font=footnotesize, textfont=footnotesize}
  - \fancyfoot{} # cues the use of fancy footer
  - \fancyhead[L]{Emilio M Bruna} # month & year on header left 
  - \fancyhead[R]{`r format(Sys.time(), '%d %B %Y')`} # name and page on header right
#Required for specifying captions t
  # - \usepackage{indentfirst}
  # - \setlength\parindent{22pt} # uncomment if you want indented paragraphs


# https://pubs.usgs.gov/search?q=&startYear=2019&__ncforminfo=Rf6y9tBB_4xn4sDOJ7tdSPmxI5Eu33dcCKy8YpAAAQRE-JJvXyAJAosPJMziEEMnxQE_iCi7dgGWVwrbQe9Gpb9a-7115VxO

---
# Publications by US Federal Personnel (Jan 2019 - May 2025)

\bigskip

```{r setup, include=FALSE}

library(tidyverse)
library(janitor)
library(gghighlight)
library(kableExtra)
# library(knitr)

knitr::opts_chunk$set(echo = FALSE, message=FALSE,warning=FALSE)

options(scipen=999)

comma <- function(x) format(x, digits = 2, big.mark = ",")

knitr::opts_chunk$set(fig.pos = 'H')

```





```{r data }

# Read data
#  fed affils -------------------------------------------------------------
# 
papers_df  <- read_rds("./data_clean/papers_df_clean.rds") %>% 
  mutate(PM=
           case_when(
             is.na(PM) ~ sample(c(1:12), 1, replace = TRUE),
             .default = as.numeric(PM)
             )
         )



authors_df  <- read_rds("./data_clean/authors_df_clean.rds") 
# 
# authors_df %>% filter(agency=="interior") %>% select(AF,authorID) %>% group_by(AF) %>% tally() %>% arrange(desc(n))


# affils_df  <- readRDS("./data_clean/affils_df_clean.rds") 
```

```{r totals }

total_pubs<-papers_df %>% 
  summarize(n=n_distinct(refID))

# Total_authors (fed+non)
total_authors<-authors_df %>% 
  select(authorID) %>% 
  distinct() %>% 
  tally()


total_federals<-authors_df %>% 
  filter(federal==TRUE) %>% 
  select(authorID) %>% 
  distinct() %>% 
  tally()

total_NOTfederals<-authors_df %>% 
  filter(federal==FALSE) %>% 
  select(authorID) %>% 
  distinct() %>% 
  tally() 
  

first_authors <- authors_df %>%
  filter(federal == TRUE) %>%
  filter(author_order == 1) 


prop_papers_fed_1st<-nrow(first_authors)/ total_pubs*100

last_authors <- authors_df %>%
  group_by(refID) %>%
  slice_tail() %>%
  filter(author_order != 1) %>%
  filter(federal == TRUE) 

prop_papers_fed_last<-nrow(last_authors)/ total_pubs*100


all_author_positions <- authors_df %>%
  filter(federal == TRUE) %>%
  distinct(refID,agency,.keep_all=TRUE) %>% 
  arrange(refID)

```



## Summary: All Agencies

### Total papers (2019-2024)

\bigskip

- Total N = `r comma(total_pubs$n)`

### Total authors (2019-2025 to date)

\bigskip

- Total N = `r comma(total_authors$n)`

- Federal Agency Authors: N = `r comma(total_federals$n)`

- Non-Federal Agency Authors: N = `r comma(total_NOTfederals$n)`


### Author Position (2019-2025)
\bigskip

- Articles with Federal Author in 1st Author Position:  `r comma(prop_papers_fed_1st)`%

 - Articles with Federal Author in Last Author position:  `r comma(prop_papers_fed_last)`%

\bigskip
\bigskip

### Number of authors per publication

\bigskip

```{r authors_per_pub }
# Authors per pub

auth_per_pub<-authors_df %>% 
  group_by(refID) %>% 
summarize(fed=sum(federal==TRUE),
          nonFed=sum(federal==FALSE),
          total=sum(fed+nonFed))

auth_per_pub_means<-auth_per_pub %>% 
  ungroup() %>% 
  drop_na() %>% 
  summarize(
    avg_Fed=mean(fed),
    sd_Fed=sd(fed),
    avg_NonFed=mean(nonFed),
    sd_NonFed=sd(nonFed),
    avg_Total=mean(total),
    sd_Total=sd(total)
    ) %>%  
 pivot_longer(
    cols = starts_with("avg_"),
    names_to = "author_category",
    names_prefix = "avg_",
    values_to = "mean_per_pub",
    values_drop_na = TRUE
  ) %>% 
  mutate(sd=if_else(author_category=="Total",sd_Total,NA)) %>% 
  mutate(sd=if_else(author_category=="NonFed",sd_NonFed,sd)) %>% 
  mutate(sd=if_else(author_category=="Fed",sd_Fed,sd)) %>% 
  select(-sd_Fed,-sd_NonFed,-sd_Total) %>% 
  rename(SD=sd,
         `Mean per Paper`=mean_per_pub,
         `Author Category`=author_category)
  

auth_per_pub_means %>%
kable(digits = 2,
  format = "latex",
  align = "lcc",
  caption = "Avg No of Federal and NonFederal Authors Per Publication",
  escape = FALSE,
  row.names = FALSE,
  booktabs = T,
  linesep = ""
) %>%
  kable_styling(
    bootstrap_options = c("hover"),
    # full_width = F,
    latex_options = c("scale_down","hold_position"),
    font_size = 12,
    position = "center"
  )

```



<!-- ## Journals -->

```{r journals }

journals <- papers_df %>%
  select(refID, DI,SO, PY, PM) %>%
  mutate_all(tolower)


all_jrnls <- journals %>%
  group_by(SO) %>%
  tally() %>%
  arrange(desc(n)) 


top_X_jrnls <- journals %>%
  group_by(SO) %>%
  tally() %>%
  arrange(desc(n)) %>%
  slice_head(n = 20)



top_N_jrnls_peryr <-
  journals %>%
  group_by(PY, SO) %>%
  tally() %>%
  arrange(desc(PY), desc(n)) %>%
  group_by(PY) %>%
  slice_head(n = 5)

```



### Total papers by year (2019-2025 to date)

```{r pubs_year, fig.height = 4,fig.width=4}

# papers by yr
pubs_yr <- papers_df %>%
  group_by(PY) %>%
  tally() %>%
  arrange(PY)

# pubs_yr_tbl<-
# pubs_yr %>% 
# kable(digits = 2,
#   format = "latex",
#   align = "ll",
#   escape = FALSE,
#   row.names = FALSE,
#   booktabs = T,
#   linesep = ""
# ) %>% 
#   kable_styling(
#     bootstrap_options = c("hover"),
#     # full_width = F,
#     latex_options = c("scale_down"),
#     latex_options = c("hold_position"),
#     font_size = 12,
#     position = "center"
#   )

pubs_yr %>% 
  filter(PY<2025) %>% 
  ggplot(aes(x=PY, y=n)) +
  geom_bar(stat="identity")+
  expand_limits(y = 0)+
  theme_classic()+
  scale_y_continuous(expand = c(0, 0), breaks=seq(0,85000,by=5000))+
  scale_x_continuous( seq(2019, 2024, by=1))
# +
#   gghighlight(PY == 2025)
```



### Articles Published Monthly (January-December 2019-2024)

\bigskip

```{r pubs_monthly1, fig.align='center',fig.cap="Articles per month (2019-2024).",fig.height = 4,fig.width=4}

### Papers per month per year 2019-2024
pubs_mo_yr <-
  papers_df %>%
  group_by(PM, PY) %>%
  tally() %>%
  mutate(PM=as.numeric(PM),
         PY=as.numeric(PY)) %>% 
  arrange(PY, PM) %>% 
  ungroup() %>% 
  mutate(month=row_number())

pubs_per_quarter<-pubs_mo_yr %>% 
  mutate(PM=as.numeric(PM)) %>% 
  mutate(Q=cut(PM, breaks = c(0, 3, 6, 9,12), 
               labels = c("Q1", "Q2", "Q3", "Q4"))) %>% 
  group_by(PY,Q) %>% 
  summarize(n=sum(n))

pubs_mo_yr %>% 
  filter(PY<2025) %>% 
  ggplot(aes(x=PM, y=n,group=PY,color=PY)) +
  geom_line() + 
  geom_point()+
  expand_limits(y = 0)+
  theme_classic()+
  scale_x_continuous( breaks=seq(1,12,by=1))+
  scale_y_continuous(expand = c(0, 0), breaks=seq(0,(max(pubs_mo_yr %>% select(n))+5000),by=2500))+
  # scale_y_continuous(expand = c(0, 0), n.breaks = 20, limits = c(0, max(pubs_mo_yr %>% select(n))+500))+
  # gghighlight(min(n) < 50)
  gghighlight(PY == 2024)
```

\bigskip
\bigskip

```{r pubs_monthly2, fig.align='center',fig.cap="Articles per Quarter (2019-2024).",fig.height = 4,fig.width=4}

pubs_per_quarter %>% 
  filter(PY<2025) %>% 
  ggplot(aes(x=Q, y=n,group=PY,color=PY)) +
  geom_line() + 
  geom_point()+
  expand_limits(y = 0)+
  theme_classic()+
  # scale_y_continuous(expand = c(0, 0), n.breaks = 4, limits = c(1, ))+
  scale_y_continuous(expand = c(0, 0), breaks=seq(0,(max(pubs_per_quarter %>% select(n))+5000),by=2500))+
  # scale_y_continuous(expand = c(0, 0), n.breaks = 20, limits = c(0, max(pubs_per_quarter %>% select(n))+500))+
  # gghighlight(min(n) < 50)
  gghighlight(PY == 2024)

```





### Monthly publications from January 1 to May 31 (2019-2025)

\bigskip

```{r fig_monthly_compare, fig.align='center',fig.cap="Monthly publications from January 1 to May 31 (2019-2025).",fig.height = 4,fig.width=4}


pubs_mo_yr %>%
  # mutate(PM=if_else(PY==2025,5,PM)) %>% ######## THIS IS KEY
  filter(PM<6) %>% 
  ggplot(aes(x=PM, y=n,group=PY,color=PY)) +
  geom_line() + 
  geom_point()+
  expand_limits(y = 0)+
  theme_classic()+
  scale_x_continuous( breaks=seq(1,5,by=1))+
  scale_y_continuous(expand = c(0, 0), breaks = seq(0, max(pubs_mo_yr %>% select(n))+5000,by=2500))+
  # gghighlight(min(n) < 50)
  gghighlight(PY == 2025)

```

### Total Publications between January 1 - May 31 (2019-2025)

\bigskip

```{r fig_total_to_date1, fig.align='center',fig.cap="Total Publications between January 1 - May 31 (2019-2025).",fig.height = 4,fig.width=6}

total_jan_may<-pubs_mo_yr %>%
  # mutate(PM=if_else(PY==2025,5,PM)) %>% ######## THIS IS KEY
  filter(PM<6) %>% 
  group_by(PY) %>% 
  summarize(n=sum(n)) %>% 
  mutate(n_diff=n-lag(n)) %>% 
  mutate(perc_previous_yr=n_diff/lag(n)*100)


# n_2024<-total_jan_may %>% filter(PY==2024) %>% select(n)
# n_2025<-total_jan_may %>% filter(PY==2025) %>% select(n)
# perc_2425<-as.numeric((n_2025-n_2024)/n_2024*100)
# diff_2425<-n_2025-n_2024


total_jan_may %>% 
  ggplot(aes(x=PY, y=n)) +
  geom_bar(stat="identity")+
  expand_limits(y = 0)+
  theme_classic()+
  annotate(geom="text", x=2024, y=(max(total_jan_may$n)-.02*max(total_jan_may$n)), label=(total_jan_may %>% filter(PY==2024) %>% select(n)),
                 color="navyblue")+
  annotate(geom="text", x=2025, y=(max(total_jan_may$n)-.02*max(total_jan_may$n)), label=(total_jan_may %>% filter(PY==2025) %>% select(n)),
           color="navyblue")+
  
  annotate(geom="text", x=2025, y=(max(total_jan_may$n)-.13*max(total_jan_may$n)), label=paste("(", round((total_jan_may %>% filter(PY==2025) %>% select(perc_previous_yr)),2),"%)",sep=""),
           color="red", size=4)+
  scale_y_continuous(expand = c(0, 0), breaks=seq(0,(max(total_jan_may %>% select(n))+500),by=2000))+
  # scale_y_continuous(expand = c(0, 0), n.breaks = 20, limits = c(0, max(total_jan_may %>% select(n))+500))+
  scale_x_continuous( breaks=seq(2019,2025,by=1))+
  gghighlight(PY == 2025)
```



```{r fig_total_to_date2, fig.align='center',fig.cap="Percent change in articles published between Jan-May relative to the same period the previous year.",fig.height = 4,fig.width=4}
# 
# total_jan_may %>% 
#   drop_na() %>% 
#   ggplot(aes(x=PY, y=perc_previous_yr)) +
#   geom_bar(stat="identity")+
#   expand_limits(y = 0)+
#   theme_classic()+
#   annotate(geom="text", x=2025, y=(max(total_jan_may$perc_previous_yr,na.rm = TRUE)-.2*max(total_jan_may$perc_previous_yr)), label=paste("(", round((total_jan_may %>% filter(PY==2025) %>% select(perc_previous_yr)),2),"%)",sep=""),
#            color="red", size=4)+
#   scale_y_continuous(expand = c(0, 0), n.breaks = 30, limits = c((min(total_jan_may %>% select(perc_previous_yr))-10), (max(total_jan_may %>% select(perc_previous_yr))+5)))+
#   scale_y_continuous(expand = c(0, 0), breaks=seq(2019,2025,by=1))+
#   gghighlight(PY == 2025)
```


```{r fig_total_to_date3, fig.align='center',fig.cap="Articles Published Jan-May (2019-2025).",fig.height = 4,fig.width=4}

# pubs_mo_yr %>%
#   filter(month<78) %>% 
#   ggplot(aes(x = month, y = n)) +
#   geom_line() +
#   geom_point() +
#   theme_classic() +
#   scale_y_continuous(expand = c(0, 0), n.breaks = 77, limits = c(1, 77)) +
#   scale_y_continuous(expand = c(0, 0), n.breaks = 20, limits = c(0, max(pubs_mo_yr %>% select(n)) + 500))+
#   # gghighlight(min(n) < 50)
#   gghighlight(PY == 2025)


```

\newpage

## 20 Agencies producing the most Publications from 1 January 2019 to 31 May 2025.

\bigskip




```{r agency_n,, fig.align='center',fig.cap="20 Agencies producing the most Publications from 1 January 2019 to 31 May 2025",fig.height = 8,fig.width=6}

# total_pubs_per_agency <- agency_authors %>%
#   select(-n_agencies, -refID) %>%
#   filter(PY<2025) %>% 
#   group_by(PY) %>%
#   summarise(across(where(is.numeric), sum)) %>%
#   pivot_longer(!PY, names_to = "agency", values_to = "count") %>%
#   ungroup() %>% 
#   group_by(agency) %>%
#   summarize(n = sum(count)) %>%
#   arrange(n)
# pubs_per_agency

total_pubs_per_agency <- authors_df %>% 
  filter(federal==TRUE) %>% 
  mutate(agency=if_else(agency=="us department of the interior", "interior",agency)) %>% 
  mutate(agency=if_else(agency=="federal reserve system", "frs",agency)) %>% 
  mutate(agency=if_else(agency=="us department of defense", "dod",agency)) %>% 
  select(refID,agency) %>% 
  distinct() %>% 
  drop_na() %>% 
  group_by(agency) %>% 
  summarize(n=n_distinct(refID)) %>% 
  arrange(desc(n))

agencies_past_20<-total_pubs_per_agency %>% 
  select(agency) %>% slice(21:nrow(total_pubs_per_agency)) %>% 
  mutate(agency=toupper(agency)) %>%  
  mutate(agency=if_else((agency=="STATE"|
                          agency=="EDUCATION"|
                          agency=="CONGRESS"|
                          agency=="TREASURY"|
                          agency=="LABOR"|
                          agency=="OTHER"),
                        str_to_title(agency),
                        agency)
  )

# agencies_past_20[1,]<-paste("Other agencies: ",agencies_past_20[1,],sep="")
# agencies_past_20[5,]<-paste(agencies_past_20[5,],"\n",sep="")
# agencies_past_20[10,]<-paste(agencies_past_20[10,],"\n",sep="")
# agencies_past_20[15,]<-paste(agencies_past_20[15,],"\n",sep="")
# agencies_past_20[20,]<-paste(agencies_past_20[20,],"\n",sep="")

  
agencies_over_20_1<-paste(agencies_past_20$agency[1:10],collapse=",") 
agencies_over_20_2<-paste(agencies_past_20$agency[11:25],collapse=",") 


  # affils_df %>% 
  # filter(federal==TRUE) %>% 
  # group_by(agency,PY) %>% 
  # tally() %>% 
  # ungroup() %>% 
  # group_by(agency) %>% 
  # tally() %>% 
  # arrange(n)



total_pubs_per_agency %>%
  slice_head(n=20) %>% 
  mutate(agency=toupper(agency)) %>% 
  mutate(agency=if_else((agency=="INTERIOR"|
                          agency=="COMMERCE"|
                          agency=="SMITHSONIAN"|
                          agency=="COMMERCE"|
                          agency=="STATE"|
                          agency=="FEDERAL RESERVE SYSTEM"|
                          agency=="OTHER"),
                        str_to_title(agency),
                        agency)
  ) %>% 
  rename(Agency=agency,
         N=n) %>% 
kable(digits = 2,
  format = "latex",
  caption = "N = 20 agencies producing the most publications",
  align = "ll",
  escape = FALSE,
  row.names = FALSE,
  booktabs = T,
  linesep = ""
) %>%
  kable_styling(
    bootstrap_options = c("hover"),
    # full_width = F,
    latex_options = c("scale_down","hold_position"),
    font_size = 12,
    position = "center"
  ) 

# %>% 
#   add_footnote(agencies_over_20,notation = "symbol") %>% 
#   column_spec(2, width = "5em")

```

\bigskip  
\bigskip  

**_Other agencies:_** `r agencies_over_20_1`  

`r agencies_over_20_2`


### Total Publications January - May (Agencies with  more than 10,000 articles only)


```{r agency_n_decline_first1, fig.align='center',fig.cap="Percent change in publicaitons (Jan-May) relative to same period prvious year. Only agencies producing over 10000 articles between 2019-2024.",fig.height = 4,fig.width=6}

agency_subset_over10K <- total_pubs_per_agency %>%
  filter(n > 10000) %>%
  select(agency,n) %>% 
  arrange(desc(n))

agency_subset_less10K <- total_pubs_per_agency %>%
  filter(n < 10000) %>%
  select(agency,n) %>% 
  arrange(desc(n))

agency_subset<-agency_subset_over10K$agency


agency_n_decline_first <-
  first_authors %>%
  filter(agency %in% agency_subset) %>%
  # mutate(PM=if_else(PY==2025,5,PM)) %>%
  filter(PM<6) %>%
  group_by(agency, PY) %>%
  tally() %>%
  group_by(agency) %>%
  mutate(decline_n = (n - lag(n))) %>%
  mutate(perc_previous = ((decline_n) / lag(n)) * 100) %>% 
  mutate(author_position="first")

agency_n_decline_last <-
  last_authors %>%
  filter(agency %in% agency_subset) %>%
  # mutate(PM=if_else(PY==2025,5,PM)) %>%
  filter(PM<6) %>%
  group_by(agency, PY) %>%
  tally() %>%
  group_by(agency) %>%
  mutate(decline_n = (n - lag(n))) %>%
  mutate(perc_previous = ((decline_n) / lag(n)) * 100) %>% 
  mutate(author_position="last")


agency_n_decline_any <-
  all_author_positions %>%
  filter(agency %in% agency_subset) %>%
  # mutate(PM=if_else(PY==2025,5,PM)) %>%
  filter(PM<6) %>%
  group_by(agency, PY) %>%
  tally() %>%
  group_by(agency) %>%
  mutate(decline_n = (n - lag(n))) %>%
  mutate(perc_previous = ((decline_n) / lag(n)) * 100) %>% 
  mutate(author_position="any")

agency_n_decline<-bind_rows(agency_n_decline_first,agency_n_decline_last,agency_n_decline_any) %>% mutate(PY=as.numeric(PY))


agency_n_decline_sum<- agency_n_decline %>% 
  filter(author_position=="any") %>% 
  select(agency,PY,n) %>% 
  group_by(PY) %>% 
  summarize(n=sum(n)) %>% 
  mutate(n_diff=n-lag(n)) %>% 
  mutate(perc_previous_yr=n_diff/lag(n)*100)

agency_n_decline_sum %>% 
  ggplot(aes(x=PY, y=n)) +
  geom_bar(stat="identity")+
  expand_limits(y = 0)+
  theme_classic()+
  geom_hline(yintercept = 0)+
  annotate(geom="text", x=2024, y=(max(agency_n_decline_sum$n)-.02*max(agency_n_decline_sum$n)), label=(agency_n_decline_sum %>% filter(PY==2024) %>% select(n)),
                 color="navyblue")+
  annotate(geom="text", x=2025, y=(max(agency_n_decline_sum$n)-.02*max(agency_n_decline_sum$n)), label=(agency_n_decline_sum %>% filter(PY==2025) %>% select(n)),
           color="navyblue")+
  annotate(geom="text", x=2025, y=(max(agency_n_decline_sum$n)-.17*max(agency_n_decline_sum$n)), label=paste("(", round((agency_n_decline_sum %>% filter(PY>2024) %>% select(perc_previous_yr)),2),"%)",sep=""),
           color="red", size=4)+
  scale_y_continuous(expand = c(0, 0), breaks=seq(0, max(agency_n_decline_sum %>% select(n))+5000,by=2500))+
  scale_x_continuous( breaks=seq(2019,2025,by=1))+
  gghighlight(PY == 2025)



# 
# agency_n_decline %>%
#   filter(author_position=="any") %>% 
#   filter(PY>2022) %>% 
#   drop_na() %>%
#   ggplot(aes(x = PY, y = n, group = agency, color = agency)) +
#   # ggplot(aes(x = PY, y = perc_previous, group = agency, color = agency)) +
#   # ggplot(aes(x=PY,y=n, group=agency, color=agency)) +
#   geom_point() +
#   geom_line() +
#   theme_classic() +
#   facet_wrap(vars(agency), scales = "free")
# +
  # scale_y_continuous(expand = c(0, 0), breaks = c(2019, 2025)) 


compare_agency_2425 <-
  agency_n_decline %>%
  drop_na() %>%
  filter(PY > 2023)
# 
# agency_n_decline %>%
#   drop_na() %>%
#   filter(PY > 2023) %>%
#   ggplot(aes(x = PY, y = perc_previous, group = agency, color = agency)) +
#   # ggplot(aes(x=PY,y=n, group=agency, color=agency)) +
#   geom_point() +
#   geom_line() +
#   theme_classic() +
#   facet_wrap(vars(author_position))+
#   scale_y_continuous(expand = c(0, 0), breaks = c(2019, 2025)) 
# +
  # scale_y_continuous(expand = c(0, 0), limits = c(0, 2000))+
  # gghighlight((perc_previous < -.5))
```

### Publications per Agency (Agencies with > 10,000 Publications)

<!-- #### Productivity January-May Relative to same period previous year -->


```{r agency_n_decline_first2, fig.align='center',fig.cap="Percent change in articles published from previous Jan-May (2019-2024).",fig.height = 6,fig.width=4}

agency_n_decline %>%
  drop_na() %>%
  filter(author_position=="any") %>% 
  filter(PY > 2023) %>%
  ggplot(aes(x = PY, y = perc_previous, fill = PY)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_x_continuous(expand = c(0, 0), breaks = c(2024, 2025))+
  theme_classic() +
  theme(legend.position="none")+
  geom_hline(yintercept = 0) +
  facet_wrap(vars(agency),ncol = 3)+
  gghighlight(PY==2025)
```

```{r agency_n_decline_first3, fig.align='center',fig.cap="Papers Jan-May from agencies producing over 10000 articles (2019-2024).",fig.height = 4,fig.width=6}


agency_n_decline %>%
  drop_na() %>%
  filter(author_position=="any") %>% 
  filter(PY > 2023) %>%
  ggplot(aes(x = agency, y = perc_previous, fill = PY, group=agency)) +
  geom_bar(stat = "identity", position = "dodge") +
  # scale_y_continuous(expand = c(0, 0), breaks = c(2024, 2025))+
  theme_classic() +
  theme(legend.position="none")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  geom_hline(yintercept = 0) +
  facet_wrap(vars(PY))

# scale_y_continuous(expand = c(0, 0), limits = c(0, 2000))+
```





<!-- ### Agencies with < 5,000 Publications (2019-2024) -->

<!-- #### Productivity January-May Relative to Previous Year January-May (2024-2025) -->



```{r}

agency_subset<-agency_subset_less10K$agency


agency_n_decline_first <-
  first_authors %>%
  filter(agency %in% agency_subset) %>%
  # mutate(PM=if_else(PY==2025,5,PM)) %>%
  filter(PM<6) %>%
  group_by(agency, PY) %>%
  tally() %>%
  group_by(agency) %>%
  mutate(decline_n = (n - lag(n))) %>%
  mutate(perc_previous = ((decline_n) / lag(n)) * 100) %>% 
  mutate(author_position="first")

agency_n_decline_last <-
  last_authors %>%
  filter(agency %in% agency_subset) %>%
  # mutate(PM=if_else(PY==2025,5,PM)) %>%
  filter(PM<6) %>%
  group_by(agency, PY) %>%
  tally() %>%
  group_by(agency) %>%
  mutate(decline_n = (n - lag(n))) %>%
  mutate(perc_previous = ((decline_n) / lag(n)) * 100) %>% 
  mutate(author_position="last")


agency_n_decline_any <-
  all_author_positions %>%
  filter(agency %in% agency_subset) %>%
  # mutate(PM=if_else(PY==2025,5,PM)) %>%
  filter(PM<6) %>%
  group_by(agency, PY) %>%
  tally() %>%
  group_by(agency) %>%
  mutate(decline_n = (n - lag(n))) %>%
  mutate(perc_previous = ((decline_n) / lag(n)) * 100) %>% 
  mutate(author_position="any")

# scale_y_continuous(expand = c(0, 0), limits = c(0, 2000))+
```



```{r agency_n_decline1, fig.align='center',fig.cap="----- (2019-2024).",fig.height = 4,fig.width = 4}

agency_n_decline<-bind_rows(agency_n_decline_first,agency_n_decline_last,agency_n_decline_any) %>% mutate(PY=as.numeric(PY))

compare_agency_2425 <-
  agency_n_decline %>%
  drop_na() %>%
  filter(PY > 2023)
# 
# agency_n_decline %>%
#   drop_na() %>%
#   filter(PY > 2023) %>%
#   ggplot(aes(x = PY, y = perc_previous, group = agency, color = agency)) +
#   # ggplot(aes(x=PY,y=n, group=agency, color=agency)) +
#   geom_point() +
#   geom_line() +
#   theme_classic() +
#   facet_wrap(vars(author_position))+
#   scale_y_continuous(expand = c(0, 0), breaks = c(2019, 2025)) 
# +
  # scale_y_continuous(expand = c(0, 0), limits = c(0, 2000))+
  # gghighlight((perc_previous < -.5))
```

```{r agency_n_decline2, fig.align='center',fig.cap="percent change from previous year in papers produced January-May.",fig.height = 5,fig.width=7}
# 
# agency_n_decline %>%
#   drop_na() %>%
#   filter(author_position=="any") %>% 
#   filter(PY > 2023) %>%
#   ggplot(aes(x = PY, y = perc_previous, fill = PY)) +
#   geom_bar(stat = "identity", position = "dodge") +
#   scale_y_continuous(expand = c(0, 0), breaks = c(2024, 2025))+
#   theme_classic() +
#   facet_wrap(vars(agency))+
#   gghighlight(PY==2025)

```



```{r agency_n_decline3, fig.align='center',fig.cap="----- (2019-2024).",fig.height = 4,fig.width=4}


# 
# agency_n_decline %>%
#   drop_na() %>%
#   filter(author_position=="any") %>% 
#   filter(PY > 2023) %>%
#   ggplot(aes(x = agency, y = perc_previous, fill = PY, group=agency)) +
#   geom_bar(stat = "identity", position = "dodge") +
#   # scale_y_continuous(expand = c(0, 0), breaks = c(2024, 2025))+
#   theme_classic() +
#   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
#   facet_wrap(vars(PY))

```



<!-- AF-ID(60011347) USGS in Scopus -->
<!-- 2019-2025  -->
<!-- Article, Review, Editorial, Letter, Data Paper, Book Chapter -->
<!-- N=12788 -->

<!-- OG=(https://ror.org/035a68863) USGS in WOS -->
<!-- N=13866 -->
<!-- 2019-2025  -->
<!-- Article, Review, Editorial, Letter, Data Paper, Book Chapter -->
<!-- N=13886 -->


<!-- National Park Service in SCOPUS -->
<!-- AF-ID(60016933) -->
<!-- N=1772 -->

<!-- National Park Service in WOS -->
<!-- https://ror.org/044zqqy65 -->
<!-- N=1223 -->


<!-- INTERIOR In WOS -->
<!-- OG=(https://ror.org/03v0pmy70) -->
<!-- 24488 -->


<!-- https://ror.org/035a68863 -->
<!-- USGS https://ror.org/035a68863 -->
<!-- Interior: https://ror.org/03v0pmy70 -->


<!-- Interior WOS ROR ID Search -->
<!-- https://www.webofscience.com/wos/woscc/summary/29ab9bde-f922-4502-a9ef-71f0c7a4dd7a-0160dcc2fa/date-descending/1 -->
<!-- 2019-2025 -->
<!-- Article, Review Article, Editorial Material, Letter, Data Paper, Book Chapters -->
<!-- N=14022 -->
